Centralizers of higher enveloping algebras and their large $N$ limits

### 2019-05-13 frame.tex
mz's edit

Recently there has emerged a systematic generalization of the enveloping algebra $U\fg$ of a Lie algebra $\fg$ to {\em higher} enveloping algebras:
each dg Lie algebra $\fg$ has an enveloping $E_n$ algebra $\UU_n\fg$.
Building on prior work \cite{BD, AF}, Knudsen \cite{Knudsen} has developed these ideas extensively and offered useful models as factorization algebras.
Our central goal here is to document generalizations of two important results about enveloping algebras of Lie algebras: 
\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where we view $\CC^{N}$ as the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ as the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$:
\[
\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg),
\]
\end{enumerate}
but in a way that works uniformly for these higher enveloping algebras.

Our results include a cousin of these results where we vary (1) by taking the limit as $N \to \infty$ and vary (2) by studying the derived {\em centralizer} of a map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra.
(The derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.)
These notions will be reviewed and explained in the next section below,
where we also describe our results in detail.
When $n=1$, however, note that we find that \owen{state the classical thing here!}

Our main tool for this result is the Loday-Quillen-Tsygan (LQT) theorem,
which describes the large $N$ limit of Lie algebra homology of $\gl_N(A)$ for $A$ a dg algebra.
Experts will know that enveloping $E_n$ algebras can be constructed using Chevalley-Eilenberg chains
(i.e., that they are shifted versions of $\cliels$, in a precise sense),
and hence will recognize the relevance of the LQT theorem in this context.
From this perspective, our main result can be understood as an $E_n$ generalization of the LQT theorem.

We hope these constructions and ideas have interesting applications to other large $N$ phenomena,
such as in quantum field theory and random matrix theory.

\begin{rmk}
Throughout this document, we will work with differential graded vector spaces over a field $k$ of characteristic zero.
When we write ``Lie algebra,'' we mean a dg Lie algebra; if we mean an ordinary Lie algebra (such as $\sl_2(\CC)$), we say so explicitly.
Similarly, ``algebra'' means ``dg algebra'' unless stated otherwise.
Our theorems will be stated in terms of the stable $\infty$-category of such cochain complexes (one is free to work with the appropriate stable model or pretriangulated dg category, as one wishes).
Many of our constructions involve, however, explicit manipulations with standard, explicit complexes.
To minimize confusion, we try to indicate the level at which we are working but that may be left implicit in some places.
\end{rmk}

@@

\begin{thm}
\label{thm: centralizer}
Let $f: \UU_n \fg \to B$ denote a map of $E_n$ algebras, where $\UU_n \fg$ is the enveloping $E_n$ algebra of a dg Lie algebra $\fg$.
In particular, $B$ is an $E_n$ algebra in $\fg$-modules via the map~$f$.

The centralizer ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra $\clies(\fg, B)$.
\end{thm}
\greg{shall explain/remind somewhere what is the $E_n$-structure there: namely induced by $E_n$-structure of $B$ and com mutliplication in $\clies$, namely convolution between an $E_\infty$-coalgebra and an $E_n$-algebra}
By using Knudsen's results, we have the following consequence.

\begin{cor}
Let $\rho: \fg \to \fh$ be a map of dg Lie algebras, and let $\UU_n \rho$ denote the associated map of enveloping $E_n$ algebras.
When $k$ is $\RR$ or $\CC$, 
the centralizer ${\frak Z}_{E_n}(\UU_n \rho)$ is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fh \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$.
\end{cor}

An immediate corollary, by taking the identity map, is the following.

\begin{cor}
\label{thm: center}
The center of the enveloping $E_n$ algebra $\UU_n\fg$ is the $E_n$ algebra $\clies(\fg, \UU_n\fg)$.
(It has, of course, a natural $E_{n+1}$ algebra structure.)

When $k$ is $\RR$ or $\CC$, 
it is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$.
\end{cor}
\greg{shall relate this to $P_n$-center of $\fg$. that shall also (but I have to think to the proof more), at least up to formality gives the $P_{n+1}$-structure on it. }
This result generalizes the fact that $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$, since the associative algebra $U\fg$ is modeled by the locally constant one-dimensional factorization algebra that assigns $\cliels (\fg \otimes \Omega^*_c(U))$ to each open $U \subset \RR$.

Having an explicit presentation for the centralizer of a map of $E_n$ algebras allows for immediate calculations of factorization homology.

Indeed, as shown  in \cite{Knudsen}, 
the enveloping $E_n$ algebras are homotopy $O(n)$-fixed points (in other words $\mathrm{Disk}_n$-algebras in the sense of~\cite{AF}) and hence determine factorization algebras on unoriented $n$-dimensional manifolds.

Therefore, for $M$ a closed $n$-dimensional manifold, one has 
\[
\int_M {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg \otimes \Omega^*(M), \cliels( \fh \otimes \Omega^*(M))).
\]
When $M = S^n$, we thus see
\[
\int_{S^n} {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg[\epsilon_n], \cliels( \fh[\epsilon_n]))
\]
as $\Omega^*(S^n) \simeq \CC[\epsilon_n]$ where $\epsilon_n$ is a formal variable with $|\epsilon_n| = n$ and $\epsilon_n^2 = 0$.
In the case that $n=1$, this recovers a theorem of Tamarkin-Tsygan \cite{TT} on the Hochschild cohomology of differential operators in the case of a group $G$ (see Example \ref{eg: TT}).

\begin{rmk}
The preceding results are well-known to experts in factorization homology (we discussed it with several, including Francis and Knudsen),
since they are quick consequences of published work, as will be seen below.
But these statements seem not to have be documented yet in the literature, 
and so we record them here.
\end{rmk}

Our primary large $N$ result is the following, which is the novel contribution of this paper.

\begin{thm}
\label{thm: large N}
Let $A$ be a unital algebra over a field $k$ of characteristic zero. 
The map of Lie algebras
\[
u_A: \sl_\infty(k) \hookrightarrow \gl_\infty(A),
\]
induced by the unit in $A$, determines a map 
\[
\UU_nu_A: \UU_n\sl_\infty(k) \to \UU_n\gl_\infty(A).
\]
In particular, $\UU_n\gl_\infty(A)$ is an $E_n$ algebra in $\sl_\infty(k)$-representations.

The centralizer ${\frak Z}_{E_n}(\UU_nu_A)$ of this map is the $E_n$ algebra obtained by forgetting the dg commutative algebra
\[
\clies(\sl_\infty(k)) \otimes \Sym(\Cyc_*(A)[1-n])
\]
down to an $E_n$ algebra.

When $k$ is $\RR$ or $\CC$, 
it is modeled by the locally constant factorization algebra that assigns
\[
\clies(\sl_\infty(\Omega^*(U)))\otimes \Sym(\Cyc_*(A \otimes \Omega^*_c(U))[1])
\]
to each open set $U \subset \RR^n$.
\end{thm}

@@

Although \cite{Knudsen} develops an adjunction between Lie and $E_n$ algebras in any presentable stable symmetric monoidal $\infty$-category $\cC$,
we need his result only in the setting of chain complexes over a field $k$ of characteristic zero $\Mod_k$ with tensor product~$\otimes_k$.
We summarize his result in that setting as follows.

@@

\begin{thm}
There is a commuting square of adjunctions
\[
\begin{tikzcd}
\Alg_\Lie(\Mod_k) \arrow[r, bend left=20, shift left=.5ex,"\UU_n"] \arrow[dd] & \Alg_{E_n}(\Mod_k) \arrow[l, shift left=.5ex, "\oblv"] \arrow[dd]\\
&\\
\Mod_k \arrow[r, bend left=20, shift left=.5ex, "\Sigma^{1-n}"] \arrow[uu, bend left=20, shift left=.5ex, "\Free_\Lie"] & \Mod_k \arrow[l, shift left=.5ex,"\Sigma^{n-1}"] \arrow[uu, bend right=20, shift right=.5ex, "\Free_{E_n}"']
\end{tikzcd}
\]
where $\Sigma$ denotes suspension \owen{because tikzcd does weird things with ``[1-n]'' instead}.
(Curving arrows denote left adjoints.)

@@

We provide a gloss of the main notions here but refer the reader to Section 5.3 of \cite{LurieHA} and \cite{FrancisHH} for extensive and detailed discussions.

Let $\cC^\otimes$ denote a presentable stable symmetric monoidal $\infty$-category (for us, chain complexes over a field $k$ of characteristic zero with~$\otimes_k$).

\begin{dfn}
For a map $f: A \to B$ of $E_n$ algebras in $\cC$, 
the {\em centralizer} ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra in $\cC$ that is universal among those that sit in a commuting diagram
\[
\begin{tikzcd}
& {\frak Z}_{E_n}(f) \otimes A \arrow[dr] & \\
A \arrow[ur, "u"] \arrow[rr, "f"] && B
\end{tikzcd}
\]
where $u$ is induced by the identity map $1_\cC \to {\frak Z}_{E_n}(f)$.

The {\em center} of an $E_n$ algebra $A$ is the centralizer of its identity morphism $\id_A: A \to A$.
\end{dfn}

When $n=1$ and $\cC$ just ordinary vector spaces, the centralizer of $f: A \to B$ thus consists of the subspace of elements of $B$ that commute with the image $f(A)$ inside~$B$.
The center of $A$ thus recovers the traditional notion.

An alternative characterization is found in Theorem 5.3.1.30 of~\cite{LurieHA} or Proposition 6.22 in~\cite{GTZ3}.

@@

When $n=1$ and $\cC$ is chain complexes over the field $k$, 
then $\Mod^{E_1}_A(\cC)$ is equivalent to $A-A$-bimodules and hence ${\frak Z}_{E_1}(f) \simeq {\rm Hoch}^*(A,B_f)$,
where $B_f$ denotes $B$ viewed as an $A-A$-bimodule via the map~$f$.

@@

Throughout this section, let $A$ be a dg algebra over a field $k$ of characteristic zero.
We do not assume $A$ is unital; whenever unitality leads to a stronger result, we will explicitly add it as a hypothesis.
(It is possible to work with $A_\infty$ algebras rather than dg algebras, 
but it changes nothing structurally essential while adding a distracting layer of notational complexity.)

\begin{dfn}
Let $\cC(A)$ denote the cochain complex $\Cyc_*(A)[1]$, which may be viewed as an abelian Lie algebra. 
Let $\cL(A)$ denote the dg Lie algebra $\gl_\infty(A)$.
\end{dfn}

Let $\sl_\infty$ denote $\sl_\infty(k)$, the trace-zero matrices of rank infinity with only finitely many nonzero entries in~$k$.
Let $\gl_\infty$ denote $\gl_\infty(k)$, the matrices of rank infinity with only finitely many nonzero entries in~$k$.
Let $\gl_\infty(A)$ denote the matrices of rank infinity with only finitely many nonzero entries in~$A$.
There is a natural action of $\sl_\infty$ on $\gl_\infty(A)$, 
and hence on its Chevalley-Eilenberg chains $\cliels(\cL(A))$.
A key step in the LQT theorem is the construction of a quasi-isomorphism
\beqn\label{lqtco}
\lambda^{co}_A: \cliels(\cL(A))_{\sl_\infty} \xto{\cong} \Sym(\cC(A)).
\eeqn
When $A$ is unital, the inclusion of the identity $k \hookrightarrow A$ means the action of $\sl_\infty$ on $\gl_\infty(A)$ is inner, 
and hence the projection $\cliels(\cL(A)) \to \cliels(\cL(A))_{\sl_\infty}$ is a quasi-isomorphism.
The standard statement of the LQT theorem composes these two maps.

@@

\begin{cor}
There is a natural transformation
\[
\lambda: \cliels(\cL(-)) \Longrightarrow \Sym(\cC(-))
\]
between these two functors from the dg category of unital dg algebras to the dg category of dg cocommutative Hopf algebras over~$k$.
It is a quasi-isomorphism on every algebra.
\end{cor}
\greg{why not state (in a remark) this corollary and the next one as syaing this is an equivalence in the $\infty$-categories of functor ?}
For our purposes, a variant is more useful, but this variant follows by easy, formal arguments.

@@

\begin{thm}
\label{thm: variant LQT}
There is a natural transformation
\[
\lqt: \clies(\sl_\infty, \cliels(\cL(-)))  \Longrightarrow \clies(\sl_\infty) \otimes \Sym(\cC(-))
\]
between these two functors from the dg category of {\em non}unital dg algebras to the dg category of dg Hopf algebras over~$k$.
For each nonunital dg algebra $A$, the map
\beqn\label{lqtinv}
\lqt_A: \clies(\sl_\infty, \cliels(\cL(A))) \xto{\simeq} \clies(\sl_\infty) \otimes \Sym(\cC(A))
\eeqn
is a quasi-isomorphism.
\end{thm}

@@

\begin{rmk}
Loday and Procesi proved versions of the Loday-Quillen-Tsygan theorem for the Lie algebras $\mathfrak{o}_n$ and $\mathfrak{sp}_{2n}$,
in which cyclic homology of the algebra is replaced by its dihedral homology.
As nothing substantive changes in proving analogous versions of our results above, 
we do not spell out the details here.
\end{rmk}

\begin{proof}
We take for granted that the reader is familiar with the proof of the LQT theorem as articulated in \cite{LQ, Weibel, LodayCyclic}.
Our primary modification is to replace the coinvariants $\cliels(\cL(A))_{\sl_\infty}$ by the invariants $\cliels(\cL(A))^{\sl_\infty}$.
Note that there is a canonical map
\beqn\label{invtoco}
\cliels(\cL(A))^{\sl_\infty} \hookrightarrow \cliels(\cL(A)) \to \cliels(\cL(A))_{\sl_\infty},
\eeqn
and in fact this map is an isomorphism, which we now show.

To see this claim, we need to understand how $\sl_\infty$ acts on the underlying graded module of $\cliels(\cL(A))$. Note that as a $k$-module, $\cL(A) \cong \gl_\infty \otimes A$, which is a free $k$-module. 
Each $n$th tensor power thus factors as
\[
(\cL(A)[1])^{\otimes n} \cong (\gl_\infty[1])^{\otimes n} \otimes A^{\otimes n}.
\]
Taking the coinvariants under the action of the symmetric group $S_n$
(which acts diagonally on the $\gl_\infty$ factor and the $A$ factor),
we obtain
\[
\Sym^n (\gl_\infty[1]) \otimes A.
\]
But the action of $\sl_\infty$ is then wholly on the left hand factor, leaving $A$ independent.

@@

\begin{dfn}
Let $A$ be a dg algebra over $k$, and let $\cA = A \otimes \Omega^*_c$ denote the associated cosheaf of dg algebras on~$\RR^n$.
Let $\cC_A$ denote the precosheaf $\Cyc_*(\cA)[1]$ of cochain complexes on $\RR^n$,
which assigns to each open set $U$, the 1-shifted cyclic chains on $\cA(U)$.
\end{dfn}
\greg{Here we shall either recall our model for cyclic chains, or first define the $\infty$-categorical universal statement and then precise a nice model. By the universal model, I mean define cyclic as the homotopy $S^1$-fixed point for the action on Hochschild chains. This will make the construction more transparent ?}
Note that for two algebras $B$ and $B'$,
\[
\Cyc_*(B) \oplus \Cyc_*(B') \simeq \Cyc_*(B \times B') .
\] 
The corresponding statement for ordinary Hochschild homology is induced by the projections $B \leftarrow B \times B' \to B'$, and any map of algebras that induces a quasi-isomorphism on Hochschild homology induces one on cyclic homology as well.
(See, for instance, \S2.2.12 of \cite{LodayCyclic}.)
Hence, for any finite collection of disjoint opens $U_1$, \dots, $U_k$, 
we have
\[
\cC_A(U_1 \sqcup \cdots \sqcup U_k) \simeq \cC_A(U_1) \oplus \cdots \oplus \cC_A(U_k).
\]
Moreover, for any nested sequence of inclusions $V_1 \subset V_2 \subset \cdots$ with $V = \bigcup_i V_i$,
there is a quasi-isomorphism
\[
{\rm hocolim}(\cC_A(V_1) \to \cC_A(V_2) \to \cdots) \simeq \cC_A(V).
\]
(See E.2.1.1 of \cite{LodayCyclic}.)
In this sense $\cC_A$ behaves like a cosheaf, up to quasi-isomorphism. 
This property underlies the following claim.

>>>>

Recently, a generalization of the concept of the enveloping algebra of a Lie algebra has emerged: each dg Lie algebra $\fg$ has an $E_n$-enveloping algebra $\UU_n\fg$, for every $n$, so that $\UU_1\fg$ is the usual enveloping algebra $U\fg$. Building on prior work \cite{BD, AF} [Owen's thesis?], Knudsen \cite{Knudsen} developed these ideas extensively and offered a useful factorization algebra model for the $E_n$ enveloping algebra. Our central goal here is to generalize the following results about the enveloping algebra to all $n$.

\begin{enumerate}
\item[(1)] the (underived) center of $U\sl_{N+1}$ is isomorphic to $(\Sym(\CC^{N}))^{S_N}$, where $\CC^{N}$ denotes the Cartan subalgebra of $\sl_{N+1}$ and $S_N$ is the Weyl group of $SL_{N+1}$, and
\item[(2)] the derived center of the enveloping algebra $U\fg$ is given by derived invariants of the adjoint action of $\fg$ on $U\fg$. That is to say $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$.
\end{enumerate}

Our results pertains to (1) by taking $N \to \infty$, and to (2) through the derived {\em centralizer} of the map $\UU_n\sl_\infty \hookrightarrow \UU_n\gl_\infty(A)$, where $A$ is a dg algebra. Note that the usual derived center is the centralizer of the identity map $\UU_n\fg \to \UU_n\fg$.

We will review these notions and describe our results in detail in the next section. When $n=1$,  \owen{state the classical thing here!}

Our primary tool here is the Loday-Quillen-Tsygan (LQT) theorem, which describes the large $N$ limit of the Lie algebra homology of $\gl_N(A)$, for a dg associative algebra $A$. Since the $E_n$ enveloping algebra of a Lie algebra can be constructed as a locally constant factorization algebra that is made out of the shifted Chevalley-Eilenberg chains \cite{}, the relevance of the LQT theorem in this context becomes apparent. Thus, one can think of our main result as an $E_n$ generalization of the LQT theorem. We hope these ideas have exciting applications in quantum field theory and random matrix theory, as well as to other large $N$ limit phenomena.

\begin{rmk}
Throughout this document, we will work with differential graded vector spaces over a field $k$ of characteristic zero. We refer to dg Lie algebras simply as Lie algebra, and refer to those without grading and a differential, such as $\sl_2(\CC)$, as ordinary Lie algebras. Similarly, the term algebra means dg algebra unless stated otherwise.
\end{rmk}

\begin{rmk}
We state our theorems in terms of the stable $\infty$-category cochain complexes instead of the appropriate stable model or pre-triangulated dg categories. Many of our constructions, however, involve explicit manipulations of standard complexes. We indicate the level at which we are working, when it is helpful.
\end{rmk}

@@

\begin{thm}
\label{thm: centralizer}
Let $\UU_n \fg$ is the enveloping $E_n$ algebra of a dg Lie algebra $\fg$ and $f: \UU_n \fg \to B$ a map of $E_n$ algebras, which gives $B$ the structure of an $E_n$ algebra in $\fg$-modules. Then, the centralizer ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra $\clies(\fg, B)$.
\end{thm}

\greg{shall explain/remind somewhere what is the $E_n$-structure there: namely induced by $E_n$-structure of $B$ and commutative multiplication in $\clies$, namely convolution between an $E_\infty$-coalgebra and an $E_n$-algebra}

By Knudsen's results \cite{Knudsen}, we have the following consequence.

\begin{cor}
Let $k$ is $\RR$ or $\CC$, and $\UU_n \rho:  \UU_n \fg \to \UU_n \fh$ be the induced by a map of Lie algebras $\rho: \fg \to \fh$.  Then, the centralizer ${\frak Z}_{E_n}(\UU_n \rho)$ is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fh \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$, where $\Omega^*_c(U)$ denoted the commutative differential graded algebra (cdga) of compactly supported forms on $U$.
\end{cor}

When $\rho: \fg \to \fg$ is the identity map, we have the following immediate corollary.

\begin{cor}
\label{thm: center}
The center of the enveloping $E_n$ algebra $\UU_n\fg$ is the $E_n$ algebra $\clies(\fg, \UU_n\fg)$, which by Dunn's additivity theorem has in fact a natural $E_{n+1}$ algebra structure. When $k$ is $\RR$ or $\CC$, the 
center is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$.
\end{cor}
\greg{shall relate this to $P_n$-center of $\fg$. that shall also (but I have to think to the proof more), at least up to formality gives the $P_{n+1}$-structure on it. }
This result generalizes the fact that $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$, since the associative algebra $U\fg$ is modeled by the locally constant one-dimensional factorization algebra that assigns $\cliels (\fg \otimes \Omega^*_c(U))$ to each open $U \subset \RR$.

Having an explicit presentation for the centralizer of a map of $E_n$ algebras leads to a calculations of factorization homology. The enveloping $E_n$ algebras are homotopy $O(n)$-fixed points, or a $\mathrm{Disk}_n$-algebras in the sense of~\cite{AF}, and hence determine factorization algebras on un-oriented $n$-dimensional manifolds \cite{Knudsen}. Therefore, for a closed $n$-dimensional manifold $M$, one has
\[
\int_M {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg \otimes \Omega^*(M), \cliels( \fh \otimes \Omega^*(M))).
\]
Since $\Omega^*(S^n) \simeq \CC[\epsilon_n]$, where $\epsilon_n$ is a formal variable with $|\epsilon_n| = n$ and $\epsilon_n^2 = 0$, we have
\[
\int_{S^n} {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg[\epsilon_n], \cliels( \fh[\epsilon_n]))
\]

For $n=1$, this recovers a theorem of Tamarkin and Tsygan \cite{TT} on the Hochschild cohomology of differential operators on a Lie group $G$; see Example \ref{eg: TT}.

\begin{rmk}
The preceding results are quick consequences of published work and are well-known to experts we spoke to, including John Francis and Ben Knudsen. Nevertheless, we record these results so they are documented in the literature.
\end{rmk}

Our novel contribution in this paper is the following large $N$ result. Let $A$ be a unital algebra over a field $k$ of characteristic zero, and $u_A: \sl_\infty(k) \hookrightarrow \gl_\infty(A)$, the map of Lie algebras induced by the unit in $A$. The induced map $\UU_nu_A: \UU_n\sl_\infty(k) \to \UU_n\gl_\infty(A)$ makes $\UU_n\gl_\infty(A)$ is an $E_n$ algebra in $\sl_\infty(k)$-modules.

\begin{thm}
\label{thm: large N}

The centralizer ${\frak Z}_{E_n}(\UU_nu_A)$ of the map $\UU_nu_A$ is the $E_n$ algebra obtained by forgetting the dg commutative algebra
\[
\clies(\sl_\infty(k)) \otimes \Sym(\Cyc_*(A)[1-n])
\]
down to an $E_n$ algebra.

When $k$ is $\RR$ or $\CC$, this centralizer is modeled by the locally constant factorization algebra that assigns
\[
\clies(\sl_\infty(\Omega^*(U)))\otimes \Sym(\Cyc_*(A \otimes \Omega^*_c(U))[1])
\]
to each open set $U \subset \RR^n$.
\end{thm}

@@

Although \cite{Knudsen} develops an adjunction between Lie and $E_n$ algebras in any presentable symmetric monoidal stable $\infty$-category $\cC$,
we need his result only for that of chain complexes $(\Mod_k, \otimes_k)$, where $k$ is a field of characteristic zero. We summarize Knudsen's result in this setting as follows.

@@

\begin{thm}
There is a commuting square of adjunctions
\[
\begin{tikzcd}
\Alg_\Lie(\Mod_k) \arrow[r, bend left=20, shift left=.5ex,"\UU_n"] \arrow[dd] & \Alg_{E_n}(\Mod_k) \arrow[l, shift left=.5ex, "\oblv"] \arrow[dd]\\
&\\
\Mod_k \arrow[r, bend left=20, shift left=.5ex, "\Sigma^{1-n}"] \arrow[uu, bend left=20, shift left=.5ex, "\Free_\Lie"] & \Mod_k \arrow[l, shift left=.5ex,"\Sigma^{n-1}"] \arrow[uu, bend right=20, shift right=.5ex, "\Free_{E_n}"']
\end{tikzcd}
\]
where $\Sigma$ denotes suspension \owen{because tikzcd does weird things with ``[1-n]'' instead}, \mahmoud{earlier we denoted the shift by [n]. should we change that?}.
(Curving arrows denote left adjoints.)

@@

We briefly review some of the main notions and refer the reader to \cite{LurieHA} Section 5.3 and \cite{FrancisHH} for detailed discussions. Let $\cC^\otimes$ denote a presentable stable symmetric monoidal $\infty$-category (for us, chain complexes over a field $k$ of characteristic zero with~$\otimes_k$).

\begin{dfn}
For a map $f: A \to B$ of $E_n$ algebras in $\cC$, 
the {\em centralizer} ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra in $\cC$ that is universal among those that sit in a commuting diagram
\[
\begin{tikzcd}
& {\frak Z}_{E_n}(f) \otimes A \arrow[dr] & \\
A \arrow[ur, "u"] \arrow[rr, "f"] && B
\end{tikzcd}
\]
where $u$ is induced by the identity map $1_\cC \to {\frak Z}_{E_n}(f)$. The {\em center} of an $E_n$ algebra $A$ is the centralizer of its identity morphism $\id_A: A \to A$.
\end{dfn}
When $n=1$ and $\cC$ is the category of vector spaces, the centralizer of $f: A \to B$ thus consists of the subspace of elements of $B$ that commute with the image $f(A)$ inside~$B$. Thus, the center of $A$ agrees with the traditional notion. See Theorem 5.3.1.30 of~\cite{LurieHA}, or Proposition 6.22 in~\cite{GTZ3}, for an alternative characterization.

@@

When $n=1$ and $\cC$ is chain complexes over the field $k$, then $\Mod^{E_1}_A(\cC)$ is equivalent to $A\otimes A^{op}$-modules and hence ${\frak Z}_{E_1}(f) \simeq {\rm Hoch}^*(A,B_f)$,
where $B_f$ denotes $B$ viewed as an $A\otimes A^{op}$-modules via the map~$f$.

@@

Throughout this section, let $A$ be a dg algebra over a field $k$ of characteristic zero. We do not assume $A$ is unital, unless we state otherwise. Also, it is possible to work with $A_\infty$ algebras rather than dg algebras, but it changes nothing structurally essential while it adds a distracting layer of complexity. Let $\cL(A)$ denote the dg Lie algebra $\gl_\infty(A)$ and $\cC(A)$ denote the cochain complex $\Cyc_*(A)[1]$ thought of as an abelian differential graded algebra. Let $\gl_\infty(A)$ denotes the infinite matrices with only finitely many nonzero entries in~$A$. Let $\gl_\infty$ stand for $\gl_\infty(k)$ and $\sl_\infty=\sl_\infty(k)$ denote the sub Lie algebra of the elements of trace is zero. There is a natural action of $\sl_\infty$ on $\gl_\infty(A)$, and hence on its Chevalley-Eilenberg chains $\cliels(\cL(A))$. A key step in the LQT theorem is the construction of an isomorphism
\beqn\label{lqtco}
\lambda^{co}_A: \cliels(\cL(A))_{\sl_\infty} \xto{\cong} \Sym(\cC(A)).
\eeqn
When $A$ is unital, the inclusion of the identity $k \hookrightarrow A$ means the action of $\sl_\infty$ on $\gl_\infty(A)$ is inner, 
and hence the projection $\cliels(\cL(A)) \to \cliels(\cL(A))_{\sl_\infty}$ is a quasi-isomorphism.
The standard statement of the LQT theorem composes these two maps.

@@

\begin{cor}
There is a natural transformation
\[
\lambda: \cliels(\cL(-)) \Longrightarrow \Sym(\cC(-))
\]
between these two functors from the dg category of unital dg algebras to the dg category of dg cocommutative Hopf algebras over~$k$. Furthermore, the induced natural transformation on the homotopy categories is a natural isomorphism. 
\end{cor}
\greg{why not state (in a remark) this corollary and the next one as saying this is an equivalence in the $\infty$-categories of functor ?}
For our purposes, the following variant, which follows easy by formal arguments, is more useful.

@@

\begin{thm}
\label{thm: variant LQT}
There is a natural transformation
\[
\lqt: \clies(\sl_\infty, \cliels(\cL(-)))  \Longrightarrow \clies(\sl_\infty) \otimes \Sym(\cC(-))
\]
between these two functors from the dg category of {\em non-unital} dg algebras to the dg category of dg Hopf algebras over~$k$.
For each nonunital dg algebra $A$, the map
\beqn\label{lqtinv}
\lqt_A: \clies(\sl_\infty, \cliels(\cL(A))) \xto{\simeq} \clies(\sl_\infty) \otimes \Sym(\cC(A))
\eeqn
is a quasi-isomorphism.
\end{thm}

@@

\begin{rmk}
Loday and Procesi proved versions of the LQT theorem for the Lie algebras $\mathfrak{o}_n$ and $\mathfrak{sp}_{2n}$, in which cyclic homology is replaced by dihedral homology.
As nothing substantive changes in proving analogous versions of our results above, we do not spell out the details here.
\end{rmk}

\begin{proof}
We take for granted that the reader is familiar with the proof of the LQT theorem as articulated in \cite{LQ, Weibel, LodayCyclic}.
Our primary modification is to replace the coinvariants $\cliels(\cL(A))_{\sl_\infty}$ by the invariants $\cliels(\cL(A))^{\sl_\infty}$.
Note that there is a canonical map
\beqn\label{invtoco}
\cliels(\cL(A))^{\sl_\infty} \hookrightarrow \cliels(\cL(A)) \to \cliels(\cL(A))_{\sl_\infty},
\eeqn
which we now show is an isomorphism.

To see this, we need to understand how $\sl_\infty$ acts on the underlying graded module of $\cliels(\cL(A))$. Note that as a $k$-module, $\cL(A) \cong \gl_\infty \otimes A$, which is a free $k$-module. 
Each $n$th tensor power thus factors as
\[
(\cL(A)[1])^{\otimes n} \cong (\gl_\infty[1])^{\otimes n} \otimes A^{\otimes n}.
\]
Taking the coinvariants under the action of the symmetric group $S_n$, which acts diagonally on the $\gl_\infty$ factor and the $A$ factor, 
we obtain
\[
\Sym^n (\gl_\infty[1]) \otimes A.
\]
But the action of $\sl_\infty$ is then solely on the left hand factor, leaving $A$ untouched.

@@

\begin{dfn}
Let $A$ be a dg algebra over $k$, and let $\cA = A \otimes \Omega^*_c$ denote the associated cosheaf of dg algebras on~$\RR^n$.
Let $\cC_A$ denote the precosheaf $\Cyc_*(\cA)[1]$ of cochain complexes on $\RR^n$,
which assigns to each open set $U$, the 1-shifted cyclic chains on $\cA(U)$.
\end{dfn}
\greg{Here we shall either recall our model for cyclic chains, or first define the $\infty$-categorical universal statement and then precise a nice model. By the universal model, I mean define cyclic as the homotopy $S^1$-fixed point for the action on Hochschild chains. This will make the construction more transparent ?}
\mahmoud{cyclic is the homotopy quotient and negative cyclic the homotopy fixed points}
Note that for two algebras $B$ and $B'$,
\[
\Cyc_*(B) \oplus \Cyc_*(B') \simeq \Cyc_*(B \times B') .
\] 
The corresponding statement for ordinary Hochschild homology is induced by the projections $B \leftarrow B \times B' \to B'$, and any map of algebras that induces a quasi-isomorphism on Hochschild homology induces one on cyclic homology as well.
(See, for instance, \S2.2.12 of \cite{LodayCyclic}.)
Hence, for any finite collection of disjoint opens $U_1$, \dots, $U_k$, 
we have
\[
\cC_A(U_1 \sqcup \cdots \sqcup U_k) \simeq \cC_A(U_1) \oplus \cdots \oplus \cC_A(U_k).
\]
Moreover, for any nested sequence of inclusions $V_1 \subset V_2 \subset \cdots$ with $V = \bigcup_i V_i$,
there is a quasi-isomorphism
\[
{\rm hocolim}(\cC_A(V_1) \to \cC_A(V_2) \to \cdots) \simeq \cC_A(V).
\]
(See E.2.1.1 of \cite{LodayCyclic}.)
In this sense $\cC_A$ behaves like a cosheaf, up to quasi-isomorphism. 
This property underlies the following claim.
